Estimating the proportion of beneficial mutations that are not adaptive in mammals

### 2024-12-05 manuscript/main-PLOS-revision-bbl.tex
Prepare files for revision

%! BibTeX Compiler = biber
%TC:ignore
\documentclass{article}

@@

\usepackage{biorxiv}
\usepackage[backend=biber,eprint=false,isbn=false,url=false,intitle=true,style=nature,date=year]{biblatex}
\addbibresource{references.bib}

\usepackage{url}
\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{lmodern}
\usepackage{xfrac, nicefrac}
\usepackage{bm}
\usepackage{listings, enumerate, enumitem}
\usepackage[export]{adjustbox}
\usepackage{graphicx}
\usepackage{bbold}
\usepackage{pdfpages}
\pdfinclusioncopyfonts=1
\usepackage{lineno}
\usepackage{tabu}
\usepackage{hhline}
\usepackage{multicol,multirow,array}
\usepackage{etoolbox}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{orcidlink}

\newcommand{\NS}[1]{\textcolor{red}{\textbf{\emph{[NS: #1]}}}}

@@

\title{Estimating the proportion of beneficial mutations that are not adaptive in mammals}
\rhead{\scshape Beneficial yet non-adaptive mutations}

\author{
    \large
    \textbf{T. {Latrille}$^{1\dag}$\orcidlink{0000-0002-9643-4668}, J. {Joseph}$^{2\dag}$\orcidlink{0009-0002-1312-9930}, D.~A. {Hartasánchez}$^{1}$\orcidlink{0000-0003-2596-6883}, N. {Salamin}$^{1}$\orcidlink{0000-0002-3963-4954}}\\
    \normalsize
    $^{1}$Department of Computational Biology, Université de Lausanne, Lausanne, Switzerland\\
    $^{2}$Laboratoire de Biométrie et Biologie Evolutive, UMR5558, Université Lyon 1, Villeurbanne, France \\
    $^{\dag}$These authors contributed equally to this work\\
    \texttt{\href{mailto:thibault.latrille@ens-lyon.org}{thibault.latrille@ens-lyon.org}} \\
}

% The submission checklist is available at:
% https://docs.google.com/document/d/1SScaqEAgZGpeaXxpWPDh1NOLaDDLLmkGOYgaGvbY-4o/edit#
% Response to reviewers at:
% https://docs.google.com/document/d/1kzt4uf6GgmuAVHx6-LQyzFuee9biyaWoMV6RNO6lSzI/edit?usp=sharing

@@

%TC:endignore
    \begin{abstract}
        Mutations can be beneficial by bringing innovation to their bearer, allowing them to adapt to environmental change. These mutations are typically unpredictable since they respond to an unforeseen change in the environment. However, mutations can also be beneficial because they are simply restoring a state of higher fitness that was lost due to genetic drift in a stable environment. In contrast to adaptive mutations, these beneficial non-adaptive mutations can be predicted if the underlying fitness landscape is stable and known. The contribution of such non-adaptive mutations to molecular evolution has been widely neglected mainly because their detection is very challenging. We have here reconstructed protein-coding-gene fitness landscapes shared between mammals, using mutation-selection models and a multi-species alignments across 87 mammals. These fitness landscapes have allowed us to predict the fitness effect of polymorphisms found in 28 mammalian populations. Using methods that quantify selection at the population level, we have confirmed that beneficial non-adaptive mutations are indeed positively selected in extant populations. Our work confirms that deleterious substitutions are accumulating in mammals and are being reverted, generating a balance in which genomes are damaged and restored simultaneously at different loci. We observe that beneficial non-adaptive mutations represent between 15\% and 45\% of all beneficial mutations in 24 of 28 populations analyzed, suggesting that a substantial part of ongoing positive selection is not driven solely by adaptation to environmental change in mammals.
    \end{abstract}

@@

    \section*{Significance statement}
    The extent to which adaptation to changing environments is shaping genomes is a central question in molecular evolution. To quantify the rate of adaptation, population geneticists have typically used signatures of positive selection. However, mutations restoring an ancestral state of higher fitness lost by genetic drift are also positively selected, but they do not respond to a change in the environment. In this study, we have managed to distinguish beneficial mutations that are due to changing environments and those that are restoring pre-existing functions in mammals. We show that a substantial proportion of beneficial mutations cannot be interpreted as adaptive.

@@

    Mutations predicted as $\SphyNeu$ were effectively composed of a mix of neutral and selected mutations with varying \textit{precision} (36--63\%) and \textit{recall} (32--45\%) across the different populations (table~\ref{table:proba}, fig.~\ref{fig:homo-afr-results}D for humans).
    The variable proportions between populations can be explained by the effective number of individuals in the population ($\Ne$), a major driver of selection efficacy.
    Moreover, estimates of mutation rate per generation ($u$), from \textcite{bergeron_evolution_2023} and \textcite{orlando_recalibrating_2013}, and Watterson’s $\theta$ obtained from the synonymous SFS as in \textcite{achaz_frequency_2009}, allow us to obtain $\Ne$ through $\Ne=\theta/4 u)$.
    Using correlation analyses that accounted for phylogenetic relationship (see section \ref{subsec:correlation-diversity} in Materials \& Methods), we found that higher $\Ne$ was associated with a smaller proportion of nearly-neutral mutations ($r^2=0.31$, $p=0.001$, fig.~\ref{fig:diversity}A).
    This result follows the prediction of the nearly-neutral theory and suggests that in populations with higher diversity (e.g.,~\textit{Bos} or \textit{Ovis}), discrimination between beneficial and deleterious mutations is more likely to occur (figs.~S9-S12).
    Conversely, many more mutations are effectively neutral in populations with lower diversity (e.g.,~\textit{Homo}).

@@

    Additionally, we controlled that these estimations were not affected by SNP mispolarization (fig.~S19).
    We also filtered out genes under pervasive adaptation, and again found an increase in $\proba [ \SphyBen \given \SpopBen]$, consistent with our expectation (Wilcoxon signed-rank, $s=120$, $p=0.027$, table~S13).
    We assessed the impact of fitting the same functional form of DFE to the three different categories of changes $\SphyDel$, $\SphyNeu$ and $\SphyBen$.
    To this aim we computed the total amount of current selection by fitting either a single DFE on the whole dataset or by summing the other three independent DFEs.
    These disjoint estimates are well correlated, with a goodness of fit $r^2={0.95, 0.89, 0.82}$ for $\proba [\SpopDel]$, $\proba [\SpopNeu]$ and $\proba [\SpopBen]$, respectively (figs.~S20-22).
    Finally, we evaluated the effect of fitting a parametric functional form for the DFE.
    As implemented in \textcite{tataru_inference_2017}, the DFE is a mixture between a reflected gamma distribution and an exponential distribution (eq.~\ref{eq:parametric-dfe}, section~\ref{subsec:precisison_recall} in Materials \& Methods).
    Instead of using such a continuous DFE, we also tested our prediction with a non-parametric functional form for the DFE, obtaining estimates of $\proba [ \SphyBen \given \SpopBen]$ in the 8--94\% range, with a mean of 43\% (tables~S14-15).

@@

    A long-standing debate in molecular evolution is whether the variation we observe between species in protein-coding genes is primarily due to nearly-neutral mutations reaching fixation by drift or primarily due to adaptation~\cite{kimura_evolutionary_1968,ohta_nearly_1992,gillespie_substitution_1994,jensen_importance_2019}.
    Measuring the ``rate of adaptation'' in proteins, as pioneered by \textcite{mcdonald_adaptative_1991}, has been central to inform this debate~\cite{galtier_half_2024}.
    However, the McDonald \& Kreitman test detects signatures of accelerated evolution in a given terminal branch compared to an expectation based on polymorphism present in the population. It considers the fraction of substitutions that fix too quickly as ``adaptive''~\cite{mcdonald_adaptative_1991, smith_adaptive_2002, welch_estimating_2006} despite there being other processes that can lead to their fixation~\cite{galtier_adaptation_2007,rousselle_influence_2019,joseph_increased_2024a} and some of these substitutions being beneficial but non-adaptive~\cite{hartl_compensatory_1996, charlesworth_other_2007, mustonen_fitness_2009, jones_shifting_2017}.
    Here, the expectation is built on the pattern of substitutions across a phylogeny compared to the fitness effects that can be estimated from both substitutions in a terminal lineage and polymorphism in populations.
    Moreover, the goal is not to detect a fraction of beneficial substitution (i.e. ``adaptive'' substitutions for \textcite{mcdonald_adaptative_1991}), but to estimate the proportion of non-adaptive mutations among beneficial ones.

@@

    Protein-coding DNA sequence alignments in placental mammals and their corresponding gene trees come from the \href{https://www.orthomam.univ-montp2.fr}{OrthoMaM} database and were processed as in \textcite{latrille_genes_2023}.
    OrthoMaM contains a total of $116$ mammalian reference sequences in v10c~\cite{ranwez_orthomam_2007, douzery_orthomam_2014, scornavacca_orthomam_2019}.

@@

    We used the Bayesian software \textit{BayesCode} (\url{https://github.com/ThibaultLatrille/bayescode}, v1.3.1) to estimate the selection coefficients for each protein-coding gene in the mammalian dataset.
    We ran the MCMC algorithm implemented in BayesCode for $2,000$ generations as described in \textcite{latrille_genes_2023}.
    For each gene, after discarding a burn-in period of $1,000$ generations of MCMC, we obtained posterior mean estimates (over the $1,000$ generations left of MCMC) of the mutation rate matrix ($\UniDimArray{\mu}$) as well as the $20$ amino-acid fitness landscape ($\UniDimArray{F^{(i)}}$) at each site $i$.

@@

    The genetic variants representing the population level polymorphisms were obtained from the following species and their available datasets: \textit{Equus caballus} (EquCab2 assembly in the EVA study PRJEB9799~\cite{alabri_whole_2020}), \textit{Bos taurus} (UMD3.1 assembly in the NextGen project: \url{https://projects.ensembl.org/nextgen/}), \textit{Ovis aries} (Oar\_v3.1 assembly in the NextGen project), \textit{Capra hircus} (CHIR1 assembly in the NextGen project), converted to ARS1 assembly with dbSNP identifiers~\cite{sherry_dbsnp_2001}), \textit{Chlorocebus sabaeus} (ChlSab1.1 assembly in the EVA project PRJEB22989~\cite{svardal_ancient_2017}), \textit{Homo sapiens} (GRCh38 assembly in the 1000 Genomes Project~\cite{zheng-bradley_alignment_2017}).
    In total, we analyzed 28 populations across the 6 different species with polymorphism data.
    The data was processed as described in \textcite{latrille_genes_2023}.

@@

    \subsection{Correlation with effective population size ($\Ne$)}
    \label{subsec:correlation-diversity}
    Genetic diversity estimator Watterson's $\thetaSyn$ was obtained for each population from the synonymous SFS as in \textcite{achaz_frequency_2009}.
    For each popuation, $\Ne$ was estimated from the equation $\Ne=\thetaSyn / (4 \times u)$, where $u$ is the mutation rate per generation.
    Estimates for $u$ were averaged per species across the pedigree-based estimation in \textcite{bergeron_evolution_2023} for \textit{Homo}, \textit{Bos}, \textit{Capra} and \textit{Chlorocebus}.
    For \textit{Ovis} we used the estimated $u$ of \textit{Capra}.
    For \textit{Equus}, we used $u$ as estimated in \textcite{orlando_recalibrating_2013} ($u=7.24\times10^{-9}$).
    Because a correlation must account for phylogenetic relationship and non-independence of samples, we fitted a Phylogenetic Generalized Linear Model in \textit{R} with the method \texttt{pgls} with default settings from the package \texttt{caper}~\cite{orme_caper_2013}.
    The mammalian dated tree was obtained from TimeTree~\cite{kumar_timetree_2017} and pruned to include only the species analysed in this study, with multi-furcation of the different populations from each species placed at the same divergence time as the species (section S4.1).

@@

%TC:ignore
    \printbibliography
%TC:endignore
    \newpage
    \begin{figure*}[!htb]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.fitness-landscape}
        \caption{
            (A \& B) For a given codon position of a protein-coding DNA sequence, amino acids (x-axis) have different fitness values (y-axis).
            Under a changing fitness landscape (A), these fitnesses fluctuate with time.
            The protein sequence follows the moving target defined by the amino-acid fitnesses. Since substitutions are preferentially accepted if they are in the direction of this target, substitutions are, on average, adaptive.
            At the phylogenetic scale (C), beneficial substitutions are common (positive signs), promoting phenotype diversification across species.
            Under a stable fitness landscape (B), most mutations reaching fixation are either slightly deleterious reaching fixation due to drift or are beneficial non-adaptive mutations restoring a more optimal amino acid.
            At the phylogenetic scale (D), deleterious substitutions (negative signs) are often reverted via beneficial non-adaptive mutations (positive signs), promoting phenotype stability and preserving well-established biological systems.
            Even though, individually, any beneficial non-adaptive mutation might have a weak effect on its bearer, we expect them to be scattered across the genome and the genome-wide signature of beneficial non-adaptive mutations to be detectable and quantifiable.}
        \label{fig:fitness-landscape}
    \end{figure*}

    \begin{figure*}[!h]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.method.proba}
        \caption{
            Selection coefficients at the phylogenetic and population scales.
            At the phylogenetic scale (A), we estimated the amino-acid fitness for each site from protein-coding DNA alignments using mutation-selection codon models.
            For every possible mutation, the difference in amino-acid fitness before and after the mutation allows us to compute the selection coefficient at the phylogenetic scale ($\Sphy$).
            Depending on $\Sphy$ (B) mutations can be predicted as deleterious ($\SphyDel$), nearly-neutral ($\SphyNeu$) or beneficial non-adaptive mutations ($\SphyBen$) toward a fitter amino acid and repairing existing functions.
            At the population scale, each observed single nucleotide polymorphism (SNP) segregating in the population can also be classified according to its $\Sphy$ value (C).
            Occurrence and frequency in the population of non-synonymous polymorphisms, contrasted to synonymous polymorphisms (deemed neutral) is used to estimate selection coefficients (D-E) at the population scale ($\Spop$), for each class of selection ($\SphyDel$, $\SphyNeu$, $\SphyBen$).
            We can thus assess whether $\Sphy$ predicts $\Spop$ and compute \textit{precision} (F) and \textit{recall} (G) for each class.
            The \textit{recall} value for class $\SphyBen$ is the probability for beneficial mutations to be non-adaptive (G).
        }
        \label{fig:method}
    \end{figure*}

    \begin{figure*}[!h]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.homo-afr-results}
        \caption{
            (A) Distribution of scaled selection coefficients ($\Sphy$), predicted for all possible non-synonymous DNA mutations away from the ancestral human exome (section \ref{subsec:nunber-of-sites}).
            Mutations are divided into three classes of selection: deleterious ($\SphyDel$), nearly-neutral ($\SphyNeu$) and beneficial ($\SphyBen$, supposedly beneficial non-adaptive mutations)
            (B) Distribution of scaled selection coefficients ($\Sphy$) for all observed substitutions along the \textit{Homo} branch after the \textit{Homo}-\textit{Pan} split (section \ref{subsec:substitution-mapping-in-the-terminal-branch}).
            If there are fewer substitutions than expected, this class is thus undergoing purifying selection, as is the case for $\SphyDel$.
            (C) The site-frequency spectrum (SFS) in humans of African descent for a random sample of 16 alleles (means in solid lines and standard deviations in color shades) for each class of selection and for synonymous mutations, supposedly neutral (black). The SFS represents the proportion of mutations (y-axis) with a given number of derived alleles in the population (x-axis).
            At high frequencies, deleterious mutations are underrepresented.
            (D) Proportion of beneficial $\ProbaPopDel$, nearly-neutral $\ProbaPopNeu$, and deleterious mutations $\ProbaPopBen$ estimated at the population scale for each class of selection at the phylogenetic scale (section \ref{subsec:s-polymorphism-method}). Proportions depicted here are not weighted by their mutational opportunities.
        }
        \label{fig:homo-afr-results}
    \end{figure*}

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.diversity}
        \caption{
            Populations in circles, mean of the species across the populations as squares.
            (A) Proportion of nearly-neutral mutations at the population scale ($\proba [ \SpopNeu]$ in the y-axis), shown as a function of estimated effective population size ($\Ne$ in the x-axis).
            (B) Proportion of beneficial non-adaptive mutations among all beneficial mutations ($\proba [ \SphyBen \given \SpopBen]$ in the y-axis), shown as a function of $\Ne$ in x-axis.
            Correlations account for phylogenetic relationship and non-independence of samples, through the fit of a Phylogenetic Generalized Linear Model (see section~\ref{subsec:s-polymorphism-method} in Materials \& Methods).
        }
        \label{fig:diversity}
    \end{figure*}

>>>>

\documentclass{article}

@@

\renewenvironment{abstract}
{
    \centerline
    {\large \bfseries \scshape Abstract}
    \begin{quote}
    }
    {
    \end{quote}
}

\def\keywordname{{\bfseries \emph Keywords}}%
\def\keywords#1{\par\addvspace\medskipamount{\rightskip=0pt plus1cm
\def\and{\ifhmode\unskip\nobreak\fi\ $\cdot$
}\noindent\keywordname\enspace\ignorespaces#1\par}}

%\usepackage{biorxiv}

\usepackage{url}
\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{lmodern}
\usepackage{xfrac, nicefrac}
\usepackage{bm}
\usepackage{listings, enumerate, enumitem}
\usepackage[export]{adjustbox}
\usepackage{graphicx}
\usepackage{bbold}
\usepackage{pdfpages}
\pdfinclusioncopyfonts=1
\usepackage{lineno}
\usepackage{tabu}
\usepackage{hhline}
\usepackage{multicol,multirow,array}
\usepackage{etoolbox}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{orcidlink}
\usepackage[verbose=true,letterpaper]{geometry}
\AtBeginDocument{
    \newgeometry{
        textheight=9in,
        textwidth=6.5in,
        top=1in,
        headheight=14pt,
        headsep=25pt,
        footskip=30pt
    }
}

\bibliographystyle{plos2015}

@@

\title{Estimating the proportion of beneficial mutations that are not adaptive in mammals}

\author{
    \small \textbf{T. {Latrille}$^{1\dag}$\orcidlink{0000-0002-9643-4668}, J. {Joseph}$^{2\dag}$\orcidlink{0009-0002-1312-9930}, D.~A. {Hartasánchez}$^{1}$\orcidlink{0000-0003-2596-6883}, N. {Salamin}$^{1}$\orcidlink{0000-0002-3963-4954}}\\
    \small $^{1}$Department of Computational Biology, Université de Lausanne, Lausanne, Switzerland\\
    \small $^{2}$Laboratoire de Biométrie et Biologie Evolutive, UMR5558, Université Lyon 1, Villeurbanne, France \\
    \small $^{\dag}$These authors contributed equally to this work\\
    \small  \texttt{\href{mailto:thibault.latrille@ens-lyon.org}{thibault.latrille@ens-lyon.org}} \\
}

@@

    \begin{abstract}
        Mutations can be beneficial by bringing innovation to their bearer, allowing them to adapt to environmental change. These mutations are typically unpredictable since they respond to an unforeseen change in the environment. However, mutations can also be beneficial because they are simply restoring a state of higher fitness that was lost due to genetic drift in a stable environment. In contrast to adaptive mutations, these beneficial non-adaptive mutations can be predicted if the underlying fitness landscape is stable and known. The contribution of such non-adaptive mutations to molecular evolution has been widely neglected mainly because their detection is very challenging. We have here reconstructed protein-coding-gene fitness landscapes shared between mammals, using mutation-selection models and a multi-species alignments across 87 mammals. These fitness landscapes have allowed us to predict the fitness effect of polymorphisms found in 28 mammalian populations. Using methods that quantify selection at the population level, we have confirmed that beneficial non-adaptive mutations are indeed positively selected in extant populations. Our work confirms that deleterious substitutions are accumulating in mammals and are being reverted, generating a balance in which genomes are damaged and restored simultaneously at different loci. We observe that beneficial non-adaptive mutations represent between 15\% and 45\% of all beneficial mutations in 24 of 28 populations analyzed, suggesting that a substantial part of ongoing positive selection is not driven solely by adaptation to environmental change in mammals.
    \end{abstract}

@@

    \section*{Author summary}
    The extent to which adaptation to changing environments is shaping genomes is a central question in molecular evolution. To quantify the rate of adaptation, population geneticists have typically used signatures of positive selection. However, mutations restoring an ancestral state of higher fitness lost by genetic drift are also positively selected, but they do not respond to a change in the environment. In this study, we have managed to distinguish beneficial mutations that are due to changing environments and those that are restoring pre-existing functions in mammals. We show that a substantial proportion of beneficial mutations cannot be interpreted as adaptive.

@@

    Mutations predicted as $\SphyNeu$ were effectively composed of a mix of neutral and selected mutations with varying \textit{precision} (36--63\%) and \textit{recall} (32--45\%) across the different populations (table~\ref{table:proba}, fig.~\ref{fig:homo-afr-results}D for humans).
    The variable proportions between populations can be explained by the effective number of individuals in the population ($\Ne$), a major driver of selection efficacy.
    Moreover, estimates of mutation rate per generation ($u$), from Bergeron \textit{et al.}~\cite{bergeron_evolution_2023} and Orlando \textit{et al.}~\cite{orlando_recalibrating_2013}, and Watterson’s $\theta$ obtained from the synonymous SFS as in Achaz~\cite{achaz_frequency_2009}, allow us to obtain $\Ne$ through $\Ne=\theta/4 u)$.
    Using correlation analyses that accounted for phylogenetic relationship (see section \ref{subsec:correlation-diversity} in Materials \& Methods), we found that higher $\Ne$ was associated with a smaller proportion of nearly-neutral mutations ($r^2=0.31$, $p=0.001$, fig.~\ref{fig:diversity}A).
    This result follows the prediction of the nearly-neutral theory and suggests that in populations with higher diversity (e.g.,~\textit{Bos} or \textit{Ovis}), discrimination between beneficial and deleterious mutations is more likely to occur (figs.~S9-S12).
    Conversely, many more mutations are effectively neutral in populations with lower diversity (e.g.,~\textit{Homo}).

@@

    Additionally, we controlled that these estimations were not affected by SNP mispolarization (fig.~S19).
    We also filtered out genes under pervasive adaptation, and again found an increase in $\proba [ \SphyBen \given \SpopBen]$, consistent with our expectation (Wilcoxon signed-rank, $s=120$, $p=0.027$, table~S13).
    We assessed the impact of fitting the same functional form of DFE to the three different categories of changes $\SphyDel$, $\SphyNeu$ and $\SphyBen$.
    To this aim we computed the total amount of current selection by fitting either a single DFE on the whole dataset or by summing the other three independent DFEs.
    These disjoint estimates are well correlated, with a goodness of fit $r^2={0.95, 0.89, 0.82}$ for $\proba [\SpopDel]$, $\proba [\SpopNeu]$ and $\proba [\SpopBen]$, respectively (figs.~S20-22).
    Finally, we evaluated the effect of fitting a parametric functional form for the DFE.
    As implemented in Tataru \textit{et al.}~\cite{tataru_inference_2017}, the DFE is a mixture between a reflected gamma distribution and an exponential distribution (eq.~\ref{eq:parametric-dfe}, section~\ref{subsec:precisison_recall} in Materials \& Methods).
    Instead of using such a continuous DFE, we also tested our prediction with a non-parametric functional form for the DFE, obtaining estimates of $\proba [ \SphyBen \given \SpopBen]$ in the 8--94\% range, with a mean of 43\% (tables~S14-15).

@@

    A long-standing debate in molecular evolution is whether the variation we observe between species in protein-coding genes is primarily due to nearly-neutral mutations reaching fixation by drift or primarily due to adaptation~\cite{kimura_evolutionary_1968,ohta_nearly_1992,gillespie_substitution_1994,jensen_importance_2019}.
    Measuring the ``rate of adaptation'' in proteins, as pioneered by McDonald \& Kreitman~\cite{mcdonald_adaptative_1991}, has been central to inform this debate~\cite{galtier_half_2024}.
    However, the McDonald \& Kreitman test detects signatures of accelerated evolution in a given terminal branch compared to an expectation based on polymorphism present in the population. It considers the fraction of substitutions that fix too quickly as ``adaptive''~\cite{mcdonald_adaptative_1991, smith_adaptive_2002, welch_estimating_2006} despite there being other processes that can lead to their fixation~\cite{galtier_adaptation_2007,rousselle_influence_2019,joseph_increased_2024a} and some of these substitutions being beneficial but non-adaptive~\cite{hartl_compensatory_1996, charlesworth_other_2007, mustonen_fitness_2009, jones_shifting_2017}.
    Here, the expectation is built on the pattern of substitutions across a phylogeny compared to the fitness effects that can be estimated from both substitutions in a terminal lineage and polymorphism in populations.
    Moreover, the goal is not to detect a fraction of beneficial substitution (i.e. ``adaptive'' substitutions for McDonald \& Kreitman~\cite{mcdonald_adaptative_1991}), but to estimate the proportion of non-adaptive mutations among beneficial ones.

@@

    Protein-coding DNA sequence alignments in placental mammals and their corresponding gene trees come from the \href{https://www.orthomam.univ-montp2.fr}{OrthoMaM} database and were processed as in Latrille \textit{et al.}~\cite{latrille_genes_2023}.
    OrthoMaM contains a total of $116$ mammalian reference sequences in v10c~\cite{ranwez_orthomam_2007, douzery_orthomam_2014, scornavacca_orthomam_2019}.

@@

    We used the Bayesian software \textit{BayesCode} (\url{https://github.com/ThibaultLatrille/bayescode}, v1.3.1) to estimate the selection coefficients for each protein-coding gene in the mammalian dataset.
    We ran the MCMC algorithm implemented in BayesCode for $2,000$ generations as described in Latrille \textit{et al.}~\cite{latrille_genes_2023}.
    For each gene, after discarding a burn-in period of $1,000$ generations of MCMC, we obtained posterior mean estimates (over the $1,000$ generations left of MCMC) of the mutation rate matrix ($\UniDimArray{\mu}$) as well as the $20$ amino-acid fitness landscape ($\UniDimArray{F^{(i)}}$) at each site $i$.

@@

    The genetic variants representing the population level polymorphisms were obtained from the following species and their available datasets: \textit{Equus caballus} (EquCab2 assembly in the EVA study PRJEB9799~\cite{alabri_whole_2020}), \textit{Bos taurus} (UMD3.1 assembly in the NextGen project: \url{https://projects.ensembl.org/nextgen/}), \textit{Ovis aries} (Oar\_v3.1 assembly in the NextGen project), \textit{Capra hircus} (CHIR1 assembly in the NextGen project), converted to ARS1 assembly with dbSNP identifiers~\cite{sherry_dbsnp_2001}), \textit{Chlorocebus sabaeus} (ChlSab1.1 assembly in the EVA project PRJEB22989~\cite{svardal_ancient_2017}), \textit{Homo sapiens} (GRCh38 assembly in the 1000 Genomes Project~\cite{zheng-bradley_alignment_2017}).
    In total, we analyzed 28 populations across the 6 different species with polymorphism data.
    The data was processed as described in Latrille \textit{et al.}~\cite{latrille_genes_2023}.

@@

    \subsection{Correlation with effective population size ($\Ne$)}
    \label{subsec:correlation-diversity}
    Genetic diversity estimator Watterson's $\thetaSyn$ was obtained for each population from the synonymous SFS as in Achaz~\cite{achaz_frequency_2009}.
    For each popuation, $\Ne$ was estimated from the equation $\Ne=\thetaSyn / (4 \times u)$, where $u$ is the mutation rate per generation.
    Estimates for $u$ were averaged per species across the pedigree-based estimation in~\cite{bergeron_evolution_2023} for \textit{Homo}, \textit{Bos}, \textit{Capra} and \textit{Chlorocebus}.
    For \textit{Ovis} we used the estimated $u$ of \textit{Capra}.
    For \textit{Equus}, we used $u$ as estimated in Orlando \textit{et al.}~\cite{orlando_recalibrating_2013} ($u=7.24\times10^{-9}$).
    Because a correlation must account for phylogenetic relationship and non-independence of samples, we fitted a Phylogenetic Generalized Linear Model in \textit{R} with the method \texttt{pgls} with default settings from the package \texttt{caper}~\cite{orme_caper_2013}.
    The mammalian dated tree was obtained from TimeTree~\cite{kumar_timetree_2017} and pruned to include only the species analysed in this study, with multi-furcation of the different populations from each species placed at the same divergence time as the species (section S4.1).

@@

    \begin{thebibliography}{10}

        \bibitem{darwin_origin_1859}
        Darwin C.
        \newblock On the {{Origin}} of {{Species}} by {{Means}} of {{Natural
        Selection}}, or the {{Preservation}} of {{Favoured Races}} in the
            {{Struggle}} for {{Life}}. vol. 220.
        \newblock John Murray; 1859.

        \bibitem{merrell_adaptive_1994}
        Merrell DJ.
        \newblock The {{Adaptive Seascape}}: {{The Mechanism}} of {{Evolution}}.
        \newblock U of Minnesota Press; 1994.

        \bibitem{gavrilets_adaptive_2009}
        Gavrilets S, Losos JB.
        \newblock Adaptive {{Radiation}}: {{Contrasting Theory}} with {{Data}}.
        \newblock Science. 2009;323(5915):732--737.
        \newblock doi:{10.1126/science.1157966}.

        \bibitem{enard_viruses_2016}
        Enard D, Cai L, Gwennap C, Petrov DA.
        \newblock Viruses Are a Dominant Driver of Protein Adaptation in Mammals.
        \newblock eLife. 2016;5:e12469.
        \newblock doi:{10.7554/eLife.12469}.

        \bibitem{mcdonald_adaptative_1991}
        McDonald JH, Kreitman M.
        \newblock Adaptative Protein Evolution at {{Adh}} Locus in {{Drosophila}}.
        \newblock Nature. 1991;351(6328):652--654.
        \newblock doi:{10.1038/351652a0}.

        \bibitem{smith_adaptive_2002}
        Smith NGC, {Eyre-Walker} A.
        \newblock Adaptive Protein Evolution in {{Drosophila}}.
        \newblock Nature. 2002;415(6875):1022--1024.
        \newblock doi:{10.1038/4151022a}.

        \bibitem{welch_estimating_2006}
        Welch JJ.
        \newblock Estimating the {{Genomewide Rate}} of {{Adaptive Protein Evolution}}
        in {{Drosophila}}.
        \newblock Genetics. 2006;173(2):821--837.
        \newblock doi:{10.1534/genetics.106.056911}.

        \bibitem{yang_statistical_2000}
        Yang Z, Bielawski JR.
        \newblock Statistical Methods for Detecting Molecular Adaptation.
        \newblock Trends in Ecology and Evolution. 2000;15(12):496--503.
        \newblock doi:{10.1016/S0169-5347(00)01994-7}.

        \bibitem{eyre-walker_genomic_2006}
        {Eyre-Walker} A.
        \newblock The Genomic Rate of Adaptive Evolution.
        \newblock Trends in Ecology \& Evolution. 2006;21(10):569--575.
        \newblock doi:{10.1016/j.tree.2006.06.015}.

        \bibitem{moutinho_variation_2019}
        Moutinho AF, Bataillon T, Dutheil JY.
        \newblock Variation of the Adaptive Substitution Rate between Species and
        within Genomes.
        \newblock Evolutionary Ecology. 2019;34(3):315--338.
        \newblock doi:{10.1007/s10682-019-10026-z}.

        \bibitem{lynch_mutation_2023}
        Lynch M.
        \newblock Mutation Pressure, Drift, and the Pace of Molecular Coevolution.
        \newblock Proceedings of the National Academy of Sciences.
        2023;120(27):e2306741120.
        \newblock doi:{10.1073/pnas.2306741120}.

        \bibitem{charlesworth_other_2007}
        Charlesworth J, {Eyre-Walker} A.
        \newblock The Other Side of the Nearly Neutral Theory, Evidence of Slightly
        Advantageous Back-Mutations.
        \newblock Proceedings of the National Academy of Sciences.
        2007;104(43):16992--16997.
        \newblock doi:{10.1073/pnas.0705456104}.

        \bibitem{mustonen_fitness_2009}
        Mustonen V, L{\"a}ssig M.
        \newblock From Fitness Landscapes to Seascapes: Non-Equilibrium Dynamics of
        Selection and Adaptation.
        \newblock Trends in genetics. 2009;25(3):111--119.
        \newblock doi:{10.1016/j.tig.2009.01.002}.

        \bibitem{jones_shifting_2017}
        Jones CT, Youssef N, Susko E, Bielawski JP.
        \newblock Shifting Balance on a Static Mutation--Selection Landscape: A Novel
        Scenario of Positive Selection.
        \newblock Molecular Biology and Evolution. 2017;34(2):391--407.
        \newblock doi:{10.1093/molbev/msw237}.

        \bibitem{ohta_nearly_1992}
        Ohta T.
        \newblock The Nearly Neutral Theory of Molecular Evolution.
        \newblock Annual Review of Ecology and Systematics. 1992;23(1992):263--286.

        \bibitem{gillespie_ohta_1995}
        Gillespie JH.
        \newblock On {{Ohta}}'s Hypothesis: {{Most}} Amino Acid Substitutions Are
        Deleterious.
        \newblock Journal of Molecular Evolution. 1995;40(1):64--69.
        \newblock doi:{10.1007/BF00166596}.

        \bibitem{hartl_compensatory_1996}
        Hartl DL, Taubes CH.
        \newblock Compensatory Nearly Neutral Mutations: {{Selection}} without
        Adaptation.
        \newblock Journal of Theoretical Biology. 1996;182(3):303--309.
        \newblock doi:{10.1006/jtbi.1996.0168}.

        \bibitem{sella_application_2005}
        Sella G, Hirsh AE.
        \newblock The Application of Statistical Physics to Evolutionary Biology.
        \newblock Proceedings of the National Academy of Sciences of the United States
        of America. 2005;102(27):9541--9546.
        \newblock doi:{10.1073/pnas.0501865102}.

        \bibitem{cvijovic_fate_2015}
        Cvijovi{\'c} I, Good BH, Jerison ER, Desai MM.
        \newblock Fate of a Mutation in a Fluctuating Environment.
        \newblock Proceedings of the National Academy of Sciences.
        2015;112(36):E5021--E5028.
        \newblock doi:{10.1073/pnas.1505406112}.

        \bibitem{piganeau_estimating_2003}
        Piganeau G, {Eyre-Walker} A.
        \newblock Estimating the Distribution of Fitness Effects from {{DNA}} Sequence
        Data: {{Implications}} for the Molecular Clock.
        \newblock Proceedings of the National Academy of Sciences of the United States
        of America. 2003;100(18):10335--10340.
        \newblock doi:{10.1073/pnas.1833064100}.

        \bibitem{chi_detecting_2020}
        Chi PB, Kosater WM, Liberles DA.
        \newblock Detecting {{Signatures}} of {{Positive Selection}} against a
            {{Backdrop}} of {{Compensatory Processes}}.
        \newblock Molecular Biology and Evolution. 2020;37(11):3353--3362.
        \newblock doi:{10.1093/molbev/msaa161}.

        \bibitem{bazykin_changing_2015}
        Bazykin GA.
        \newblock Changing Preferences: Deformation of Single Position Amino Acid
        Fitness Landscapes and Evolution of Proteins.
        \newblock Biology letters. 2015;11(10):20150315.
        \newblock doi:{10.1098/rsbl.2015.0315}.

        \bibitem{moses_inferring_2009}
        Moses AM, Durbin R.
        \newblock Inferring {{Selection}} on {{Amino Acid Preference}} in {{Protein
        Domains}}.
        \newblock Molecular Biology and Evolution. 2009;26(3):527--536.
        \newblock doi:{10.1093/molbev/msn286}.
